Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data

### 2020-07-02 submission/cv_models.tex
fixed fig 3
full pass in light of kdd reviewer comments

We perform \emph{coarse model analysis}, comparing and contrasting the four model series, and predicting trends in model quality. 
We also perform \emph{fine layer analysis}, as a function of depth for these models, illustrating that PL-based metrics can provide novel insights among the VGG, ResNet/ResNet-1K, and DenseNet architectures.

@@

See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, and DenseNet series.
Similar results (not shown) are obtained for the Top5 accuracies.
Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.

The DenseNet series has similar behavior to what we see in Figures~\ref{fig:vgg-metrics} and~\ref{fig:cv2-accuracy} for the other models.
However, as noted in Table~\ref{table:cv-models}, it has only 4 data points.  
In our larger analysis, in Section~\ref{sxn:all_cv_models}, we will only include series with 5 or more models.
%For completeness and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
(Note that these and many other such plots can be seen on our publicly-available~repo.)

@@

We can learn much more about a pretrained model by going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).  % in the network. 
%The most interesting results are seen when we 
For example, we can 
plot (just) the PL exponent, $\alpha$, for each layer, %
%\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
as a function of depth.
%
See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
(Again, a much more detailed set of plots is available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)

@@

    \subfigure[ VGG ]{
        \includegraphics[width=4.9cm]{img/VGG_fnl_alpha_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
                \label{fig:vgg-alpha-layers}
    }
    %KDD% 
    \qquad
    \subfigure[ ResNet-1K ]{
        \includegraphics[width=4.9cm]{img/ResNet_fnl_alpha_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
        \label{fig:resnet-alpha-layer}
    }
    %KDD% 
    \qquad
    \subfigure[ DenseNet ]{
        \includegraphics[width=4.9cm]{img/DenseNet_fnl_alpha_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
        \label{fig:densenet-alpha-layer}
    }    
    %KDD% 
    \qquad
    \subfigure[ ResNet-1K (overlaid) ]{
        \includegraphics[width=4.9cm]{img/resnet_alpha_overlaid_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
        \label{fig:resnet_alpha_overlaid_depth}
    }
    \caption{PL exponent ($\alpha$) versus layer id, for the least and the most accurate models in VGG (a), ResNet (b), and DenseNet (c) series. 
             (VGG is without BN; and note that the Y axes on each plot are different.)  
             Subfigure (d) displays the ResNet models (b), zoomed in to $\alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, from smallest to largest, to
             allow a more detailed analysis of the most strongly correlated layers.
             Notice that ResNet152 exhibits different and much more stable behavior of $\alpha$ across layers.
             This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
             In the text, this is interpreted in terms of \emph{Correlation Flow}.
            }
    \label{fig:3models-alpha-layers}
\end{figure}

@@

%KDD% 
\vspace{-1mm}
\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing \emph{residual connections}.
This optimized approach allows for greater accuracy with far fewer parameters (and GPU memory requirements), and
ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.

We conjecture that the efficiency and effectiveness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
Contrast this with the DenseNet models, which contains many connections between every layer.
Our results (large $\alpha$, meaning they even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.

@@

More generally, we can understand the results presented in Figure~\ref{fig:3models-alpha-layers} in terms of what we will call the \emph{Correlation Flow} of the model. 
Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}. 
There, one expects the weight matrices of well-trained DNNs will exhibit correlations over many size scales. 
Their ESDs can be well-fit by a (truncated) PL, with exponents $\alpha\in[2,4]$.
Much larger values $(\alpha\gg 6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
%, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.
Informally, one would expect a DNN model to perform well when it facilitates the propagation of information/features across layers.
Previous work argues this by computing the gradients over the input data.
In the absence of training/test data, one might hope that this leaves empirical signatures on weight matrices, and thus we can to try to quantify this by measuring the PL properties of weight matrices.
In this case, smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and we expect that small $\alpha$ values that are stable across multiple layers enable better \emph{correlation flow} through the network.
We have seen this in many models, including those shown in Figure~\ref{fig:3models-alpha-layers}.

@@

The similarity between norm-based metrics and PL-based metrics suggests a question: is the Weighted Alpha metric just a variation of the more familiar norm-based metrics?  %%, or does it capture something different?  
More generally, do fitted $\alpha$ values contain information not captured by norms? 
In examining hundreds of pretrained models, we have found several anomalies that demonstrate the power of our approach.
In particular, to show that $\alpha$ does capture something different, consider the following example, which looks at a compressed/distilled DNN model~\cite{CWZZ17_TR}.
In this example, we show that some distillation methods may actually \emph{break} models unexpectedly by introducing what we call \emph{Scale Collapse}, where several distilled layers have unexpectedly small Spectral Norms.

@@

The pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
VGG, ResNet, or DenseNet series.  %% we considered above.
While norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with test accuracies across different models.
We do expect, however, the PL $\alpha$ to do so because it effectively measures the amount of correlation in the model~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
The reason for the anomalous behavior shown in 
Figure~\ref{fig:resnet204D5L}
is that the \texttt{distiller} Group Regularization technique 
%%has the unusual effect of increasing
causes the norms of the $\mathbf{W}$ pre-activation maps for two Conv2D layers to increase spuriously.
This is difficult to diagnose by analyzing training/test curves, but it is easy to diagnose with our~approach.

>>>>

We first perform \emph{coarse model analysis}, comparing and contrasting the four model series, and predicting trends in model quality. 
We then perform \emph{fine layer analysis}, as a function of depth for these models.
This goes beyond simply predicting trends in model quality, instead illustrating that PL-based metrics can provide novel insights among the VGG, ResNet/ResNet-1K, and DenseNet architectures.

@@

See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, and DenseNet series.
Similar results (not shown) are obtained for the Top5 accuracies.
Overall, for the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
These model series are all very well-trodden, and thus our results indicate that norm-based metrics and PL-based metrics can both distinguish among a series of well-trained versus more-well-trained models, with PL-based metrics performing somewhat (i.e., quantitatively) better.

The DenseNet series has similar behavior to what we see in Figures~\ref{fig:vgg-metrics} and~\ref{fig:cv2-accuracy} for the other models.
However, as noted in Table~\ref{table:cv-models}, it has only 4 data points.  
In our larger analysis, in Section~\ref{sxn:all_cv_models}, we will only include series with 5 or more models.
%For completeness and reproducibility here, the DenseNet Log $\alpha$-Norm plot is given in Appendix~\ref{sxn:appendix}.
(These and many other such plots can be seen on our publicly-available~repo.)

@@

We can learn much more about a pretrained model by going beyond average values of quality metrics to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).  % in the network. 
%The most interesting results are seen when we 
For example, we can 
plot (just) the PL exponent, $\alpha$, for each layer, %
%\footnote{That is, here we consider just $\alpha$ for each layer, i.e., not $\hat{\alpha}$ or $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}$.} 
as a function of depth.
%
See Figure~\ref{fig:3models-alpha-layers}, which plots $\alpha$ for each layer (the first layer corresponds to data, the last layer to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (no BN), ResNet, and DenseNet series.
(Again, many more such plots are available at our repo; but note that the corresponding layer-wise plots for Frobenius and Spectral norms are much less interesting than the results we present here.)

@@

    \subfigure[ VGG ]{
        \includegraphics[width=4.9cm]{img/VGG_fnl_alpha_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/VGG_fnl_alpha_depth.png} 
                \label{fig:vgg-alpha-layers}
    }
    %KDD% 
    \qquad
    \subfigure[ ResNet ]{
        \includegraphics[width=4.9cm]{img/ResNet_fnl_alpha_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/ResNet_fnl_alpha_depth.png} 
        \label{fig:resnet-alpha-layer}
    }
    %KDD% 
    \qquad
    \subfigure[ DenseNet ]{
        \includegraphics[width=4.9cm]{img/DenseNet_fnl_alpha_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/DenseNet_fnl_alpha_depth.png} 
        \label{fig:densenet-alpha-layer}
    }    
    %KDD% 
    \qquad
    \subfigure[ ResNet (overlaid) ]{
        \includegraphics[width=4.9cm]{img/resnet_alpha_overlaid_depth.png} 
        %KDD% \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
        \label{fig:resnet_alpha_overlaid_depth}
    }
    \caption{PL exponent ($\alpha$) versus layer id, for the least and the most accurate models in VGG (a), ResNet (b), and DenseNet (c) series. 
             (VGG is without BN; and note that the Y axes on each plot are different.)  
             Subfigure (d) displays the ResNet models (b), zoomed in to $\alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, from smallest to largest, to
             allow a more detailed analysis of the most strongly correlated layers.
             Notice that ResNet152 exhibits different and much more stable behavior of $\alpha$ across layers.
             This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
             In the text, this is interpreted in terms of \emph{Correlation Flow}.
            }
    \label{fig:3models-alpha-layers}
\end{figure}

@@

%KDD% 
\vspace{-1mm}
\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
We can interpret these layer-wise observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing \emph{residual connections}.
This optimized approach allows for greater accuracy with far fewer parameters (and GPU memory requirements), and
ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.

We conjecture that the efficiency and effectiveness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and more strongly optimized.
Contrast this with the DenseNet models, which contains many connections between every layer.
Our results (large $\alpha$, meaning that even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly~optimized.

Fine-scale measurements such as these enable us to form hypotheses as to the inner workings of DNN models, opening the door to an improved understanding of \emph{why} DNNs work, as well as \emph{how} to design better DNN models.
We next turn to two such~examples.

@@

We can understand the results presented in Figure~\ref{fig:3models-alpha-layers} in terms of what we will call the \emph{Correlation Flow} of the model. 
Recall that the average Log $\alpha$-Norm metric and the Weighted Alpha metric are based on HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}, which is in turn based on ideas from the statistical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017}. 
There, one expects that the weight matrices of well-trained DNNs will exhibit correlations over many size scales, as is well-known in other strongly-correlated systems~\cite{BouchaudPotters03, SornetteBook}. 
In particular, this would imply that their ESDs can be well-fit by a truncated PL, with exponents $\alpha\in[2,4]$.
Much larger values $(\alpha\gg 6)$ may reflect poorer PL fits, whereas smaller values $(\alpha\sim 2)$, are associated with models that generalize better.
%, but the presence of many adjacent layers with relatively good PL fits \emph{suggests} that those adjacent layers ``impedance match'' well and that the correlations that are captured by one layer propagate well to subsequent layers.

Informally, one would expect a DNN model to perform well when it facilitates the propagation of information/features across layers.
Previous work argues this by computing the gradients over the input data.
In the absence of training/test data, one might hypothesize that this flow of information leaves empirical signatures on weight matrices, and thus that we can quantify this by measuring the PL properties of weight matrices.
In this case, given the statistical mechanical basis of HT-SR Theory, smaller $\alpha$ values would correspond to layers in which information correlations between data across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we hypothesize that small $\alpha$ values that are stable across multiple layers enable better \emph{correlation flow} through the network.
We have seen this in many models, including those shown in Figure~\ref{fig:3models-alpha-layers}.

This is superficially similar to recent work on the information bottleneck~\cite{ST17_TR}, except that here we work in an entirely unsupervised setting.
Developing theory to evaluate this hypothesis and exploit this insight is an intriguing direction for future work.

@@

The similarity between norm-based metrics and PL-based metrics suggests a question: is the Weighted Alpha metric just a variation of the more familiar norm-based metrics?  %%, or does it capture something different?  
More generally, do fitted $\alpha$ values contain information not captured by norms? 
In examining hundreds of pretrained models, we have found several anomalies that are ``exceptions that prove the rule'' and that demonstrate the power of our approach.
In particular, to show that $\alpha$ does capture something different, consider the following example, which looks at a compressed/distilled DNN model~\cite{CWZZ17_TR}.
In this example, we show that some distillation methods may actually \emph{break} models unexpectedly by introducing what we call \emph{Scale Collapse}, where several distilled layers have unexpectedly small Spectral Norms.

@@

The obvious interpretation is the following. 
Pretrained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the 
VGG, ResNet, or DenseNet series.  %% we considered above.
While norms make good regularizers for a single model, there is no reason \emph{a priori} to expect them correlate so well with test accuracies across different models, and they may not differentiate well-trained versus poorly-trained models.
We do expect, however, the PL $\alpha$ to do so, because it effectively measures the amount of information correlation in the model~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Among other things, this suggests that the $\alpha$ values will improve, i.e., decrease, over time, as distillation techniques continue to improve.
The reason for the anomalous behavior shown in 
Figure~\ref{fig:resnet204D5L}
is that the \texttt{distiller} Group Regularization technique 
%%has the unusual effect of increasing
causes the norms of the $\mathbf{W}$ pre-activation maps for two Conv2D layers to increase spuriously.
This is difficult to diagnose by analyzing training/test curves, but it is easy to diagnose with our~approach.

### 2020-07-02 submission/introduction.tex
full pass in light of kdd reviewer comments

Most importantly, in industry, one faces unique practical problems such as: do we have enough data for this model? 
Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.
Methods that are developed and evaluated on any well-defined publicly-available corpus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.  
We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.

In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and we do so at scale.
We consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
%
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
there are now hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
These provide a large corpus of models that by some community standard are state-of-the-art.%
\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.

The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrices.
They are based on norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and/or parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
Note that, while we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparameter optimization).% 
\footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here. Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask:
\begin{quote}
\emph{Can we predict trends in the quality of pretrained DNN models without access to training or testing data?}  
\end{quote}

@@

To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
Here is a summary of our main results.
\begin{itemize}
\item
\textbf{Norm-based metrics and well-trained models.}
Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
\item
\textbf{Norm-based metrics and poorly-trained models.}
Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.), exhibiting \emph{Scale Collapse} for these models. 
\item 
\textbf{PL-based metrics and model quality.}
PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
They are quantitatively better at discriminating ``good-better-best'' trends, and qualitatively better at distinguishing ``good-versus-bad''~models.
\item 
\textbf{PL-based metrics and model diagnostics.}
PL-based metrics can also be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
\end{itemize}

\noindent
We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
We have examined hundreds of models, and we identify general trends, but we also highlight interesting exceptions.
%%Several of the most interesting are described below.

>>>>

Most importantly, in industry, one faces unique practical problems such as: do we have enough data for this model? 
Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.
Methods that are developed and evaluated on any well-defined publicly-available corpus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.  
There is a need for a \emph{practical theory} for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.

In this paper, we provide a detailed empirical analysis to address these questions, evaluating quality metrics for pretrained deep neural network (DNN) models, and we do so at scale.
Our approach may be viewed as a statistical meta-analysis of previously published work, where we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
%
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
there are now hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
These publicly-available models provide a large corpus of models that by some community standard are state-of-the-art. %
%%%\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols.

In the absence of training and testing data, obvious quantities to examine are the weight matrices of pretrained models.
Thus,
the \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrices.
In particular, they are based on norms of weight matrices 
and/or parameters of power law (PL) fits of the eigenvalues of weight matrices.
Such norm-based metrics have been used in traditional statistical learning theory to bound capacity and construct regularizers; and such PL fits are based on statistical mechanics approaches to DNNs.
While we use traditional norm-based and PL-based metrics, our goals are not the traditional goals.
Unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., as with hyperparameter optimization).% 
\footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here. Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases, we ask:
\begin{quote}
\emph{Can we predict trends in the quality of pretrained DNN models without access to training or testing data?}  
\end{quote}

@@

To answer this question, we perform a statistical meta-analysis, where we 
perform weight matrix analysis on %%%analyze 
hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
Here is a summary of our main results.
\begin{itemize}
\item
\textbf{Norm-based metrics and well-trained models.}
Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
\item
\textbf{Norm-based metrics and poorly-trained models.}
Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
For example, they may exhibit what we call \emph{Scale Collapse} for these models. 
\item 
\textbf{PL-based metrics and model quality.}
PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
They are qualitatively better at discriminating well-trained versus poorly-trained models, and quantitatively better at discriminating among a series of well-trained versus more-well-trained models.
\item 
\textbf{PL-based metrics and model diagnostics.}
PL-based metrics can also be used to characterize fine-scale model properties (including what we call layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
\end{itemize}

\noindent
In light of our main results, several points are worth emphasizing.
First, our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
We have examined hundreds of models, and we identify general trends, but we also highlight interesting exceptions.
%%Several of the most interesting are described below.
Second, our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and the theory we apply was developed previously~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in machine learning, where the emphasis is on developing better training protocols.
Third, in addition to addressing our main question, our weight matrix analysis can be used to obtain fine-scale insight into DNN models.
This can be used to develop metrics for \emph{how} to train better models as well as to form hypotheses for \emph{why} DNNs perform as they do.
We provide two such examples---\emph{Scale Collapse} and \emph{Correlation Flow}---that we have observed in both CV and NLP models and that deserve attention in future work.
